---
title: "Consolidated_Plots_Code"
output: html_notebook
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(ggplot2)
library(tidyverse)
diamonds<-read_csv("diamonds4.csv")

#Check Data Loaded Properly:
head(diamonds)


```

Get size/shape of Data:

```{r}
nrow(diamonds)
ncol(diamonds)
dim(diamonds)

```

5 variables, and 1,214 rows

Checking for NA's:

```{r}
(diamonds[!complete.cases(diamonds),])
(diamonds[complete.cases(diamonds),])
```


No Na's in data-nice. 

Check for levels:

```{r}
levels(diamonds$cut)
levels(diamonds$color)
levels(diamonds$clarity)
```


No levels. We'll need to add some factors to get data in order that matches the
website, since alphabetical won't work for all data:

```{r}
diamonds<- diamonds%>%
  mutate(cut = cut%>%
           fct_relevel(c("Good","Very Good","Ideal","Astor Ideal")))

diamonds<- diamonds%>%
  mutate(clarity = clarity%>%
           fct_relevel(c("SI2","SI1","VS2","VS1", "VVS2", "VVS1", "IF", "FL")))

#Since R defaults to alphabetical order, we don't need to neccessarily add factors here, but doing so to flip order the same as above

diamonds<- diamonds%>%
  mutate(color = color%>%
           fct_relevel(c("J","I","H","G", "F", "E", "D")))

#Check new levels:

levels(diamonds$cut)
levels(diamonds$color)
levels(diamonds$clarity)

```



To visual distributions of different parameters:


```{r}
ggplot(diamonds, aes(x=cut))+
  geom_bar(fill="blue")+
  theme(plot.title = element_text(hjust = .5))+
  labs(title = "Distribution of Diamonds by Cut")

ggplot(diamonds, aes(x=clarity))+
  geom_bar(fill="red")+
  theme(plot.title = element_text(hjust = .5))+
  labs(title = "Distribution of Diamonds by Clarity")

ggplot(diamonds, aes(x=color))+
  geom_bar(fill="green")+
  theme(plot.title = element_text(hjust = .5))+
  labs(title = "Distribution of Diamonds by Color")

```



Scatter plot of Price vs Carat:

```{r}
ggplot(data = diamonds, aes(x = carat, y = price)) +
  geom_point()+
  theme(plot.title = element_text(hjust = .5))+
  labs(title = "Scatterplot of Carat against Price", 
       x = "Carat",
         y = "Price")

```

         
Scatter plot of Price vs Carat with all parameters added:
         

```{r}
ggplot(data = diamonds, aes(x = carat, y = price, color = color, size = clarity, shape = cut)) +
  geom_point()+
  theme(plot.title = element_text(hjust = .5))+
  labs(title = "Scatterplot of Price Against Carat With Other Variables", 
       x = "Carat",
       y = "Price (USD)")
```


Very Messy and not particularly useful, so let's break it down into indvidual variables:
```{r}

ggplot(data = diamonds, aes(x = carat, y = price, color = clarity))+
  geom_point()+
  theme(plot.title = element_text(hjust = .5))+
  labs(title = "Scatterplot of Price Against Carat With Clarity", 
       x = "Carat",
       y = "Price (USD)")
  
  
ggplot(data = diamonds, aes(x = carat, y = price, color = color)) +
  geom_point()+
  theme(plot.title = element_text(hjust = .5))+
  labs(title = "Scatterplot of Price Against Carat With Color", 
       x = "Carat",
       y = "Price (USD)")
  
  
ggplot(data = diamonds, aes(x = carat, y = price, color = cut)) +
  geom_point()+ 
  theme(plot.title = element_text(hjust = .5))+
  labs(title = "Scatterplot of Price Against Carat With Cut", 
                     x = "Carat",
                     y = "Price (USD)")
```

Website Claims order of importance is: Cut, color, carat, clarity


First: create new column "price/carat" to control for carat size as we explore how other attributes affect price


```{r}
diamonds<-diamonds%>%
  mutate(PricePerCarat = price/carat)
```



Get grouped means of price per carat for each attribute:

```{r}
ClarityVSPricePerCarat<-diamonds%>%
  group_by(clarity)%>%
  summarize(meanPrice=mean(PricePerCarat))

ColorVSPricePerCarat<-diamonds%>%
  group_by(color)%>%
  summarize(meanPrice=mean(PricePerCarat))

CutVSPricePerCarat<-diamonds%>%
  group_by(cut)%>%
  summarize(meanPrice=mean(PricePerCarat))
```


Plots of grouped Means:

```{r}
ggplot(ClarityVSPricePerCarat, aes(x=clarity, y=meanPrice))+
  geom_bar(stat="identity", fill = "red")+
  theme(plot.title = element_text(hjust = .5))+
  labs(title = "Mean Price per Carat by Clarity", 
       x = "Clarity",
       y = "Mean Price Per Carat")

ggplot(ColorVSPricePerCarat, aes(x=color, y=meanPrice))+
  geom_bar(stat="identity", fill = "blue")+
  theme(plot.title = element_text(hjust = .5))+
  labs(title = "Mean Price per Carat by Color", 
       x = "Color",
       y = "Mean Price Per Carat")

ggplot(CutVSPricePerCarat, aes(x=cut, y=meanPrice))+
  geom_bar(stat="identity", fill = "green")+
  theme(plot.title = element_text(hjust = .5))+
  labs(title = "Mean Price per Carat by Cut",
       x = "Cut",
       y = "Mean Price Per Carat")

```



#looking at some outliers
```{r}
diamonds%>%
  filter(clarity == "FL")

diamonds%>%
  filter(cut == "Ideal")

diamonds%>%
  filter(color == "H")
```



```{r}
ggplot(diamonds, aes(x=cut,fill=color))+
  geom_bar(position='fill')+
  theme(axis.text.x = element_text(),
        plot.title = element_text(hjust = 0.5))+
  labs(title="Distribution of cut vs color")
```
#Other than Astor Ideal (which only has 20 datapoints), good color seems to decrease as cut gets better. This corrlation could be a factor in why cut does not seem to correlate as much with price/carat.


