---
title: "Consolidated_Markdown of code"
output: html_notebook
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(ggplot2)
library(tidyverse)
diamonds<-read_csv("diamonds4.csv")

#Check Data Loaded Properly:
head(diamonds)


```

Get size/shape of Data:

```{r}
nrow(diamonds)
ncol(diamonds)
dim(diamonds)

```

5 variables, and 1,214 rows

Checking for NA's:

```{r}
(diamonds[!complete.cases(diamonds),])
(diamonds[complete.cases(diamonds),])
```


No Na's in data-nice. 

Check for levels:

```{r}
levels(diamonds$cut)
levels(diamonds$color)
levels(diamonds$clarity)
```


No levels. We'll need to add some factors to get data in order that matches the
website, since alphabetical won't work for all data:

```{r}
diamonds<- diamonds%>%
  mutate(cut = cut%>%
           fct_relevel(c("Good","Very Good","Ideal","Astor Ideal")))

diamonds<- diamonds%>%
  mutate(clarity = clarity%>%
           fct_relevel(c("SI2","SI1","VS2","VS1", "VVS2", "VVS1", "IF", "FL")))

#Since R defaults to alphabetical order, we don't need to neccessarily add factors here, but doing so to flip order the same as above

diamonds<- diamonds%>%
  mutate(color = color%>%
           fct_relevel(c("J","I","H","G", "F", "E", "D")))

#Check new levels:

levels(diamonds$cut)
levels(diamonds$color)
levels(diamonds$clarity)

```



To visual distributions of different parameters:


```{r}
ggplot(diamonds, aes(x=cut))+
  geom_bar(fill="blue")+
  theme(plot.title = element_text(hjust = .5))+
  labs(title = "Distribution of Diamonds by Cut")

ggplot(diamonds, aes(x=clarity))+
  geom_bar(fill="red")+
  theme(plot.title = element_text(hjust = .5))+
  labs(title = "Distribution of Diamonds by Clarity")

ggplot(diamonds, aes(x=color))+
  geom_bar(fill="green")+
  theme(plot.title = element_text(hjust = .5))+
  labs(title = "Distribution of Diamonds by Color")

```



Scatter plot of Price vs Carat:

```{r}
ggplot(data = diamonds, aes(x = carat, y = price)) +
  geom_point()+
  theme(plot.title = element_text(hjust = .5))+
  labs(title = "Scatterplot of Carat against Price", 
       x = "Carat",
         y = "Price")

```

         
Scatter plot of Price vs Carat with all parameters added:
         

```{r}
ggplot(data = diamonds, aes(x = carat, y = price, color = color, size = clarity, shape = cut)) +
  geom_point()+
  theme(plot.title = element_text(hjust = .5))+
  labs(title = "Scatterplot of Price Against Carat With Other Variables", 
       x = "Carat",
       y = "Price (USD)")
```


Very Messy and not particularly useful, so let's break it down into indvidual variables:
```{r}

ggplot(data = diamonds, aes(x = carat, y = price, color = clarity))+
  geom_point()+
  theme(plot.title = element_text(hjust = .5))+
  labs(title = "Scatterplot of Price Against Carat With Clarity", 
       x = "Carat",
       y = "Price (USD)")
  
  
ggplot(data = diamonds, aes(x = carat, y = price, color = color)) +
  geom_point()+
  theme(plot.title = element_text(hjust = .5))+
  labs(title = "Scatterplot of Price Against Carat With Color", 
       x = "Carat",
       y = "Price (USD)")
  
  
ggplot(data = diamonds, aes(x = carat, y = price, color = cut)) +
  geom_point()+ 
  theme(plot.title = element_text(hjust = .5))+
  labs(title = "Scatterplot of Price Against Carat With Cut", 
                     x = "Carat",
                     y = "Price (USD)")
```

Website Claims order of importance is: Cut, color, carat, clarity


First: create new column "price/carat" to control for carat size as we explore how other attributes affect price


```{r}
diamonds<-diamonds%>%
  mutate(PricePerCarat = price/carat)
```



Get grouped means of price per carat for each attribute:

```{r}
ClarityVSPricePerCarat<-diamonds%>%
  group_by(clarity)%>%
  summarize(meanPrice=mean(PricePerCarat))

ColorVSPricePerCarat<-diamonds%>%
  group_by(color)%>%
  summarize(meanPrice=mean(PricePerCarat))

CutVSPricePerCarat<-diamonds%>%
  group_by(cut)%>%
  summarize(meanPrice=mean(PricePerCarat))
```


Plots of grouped Means:

```{r}
ggplot(ClarityVSPricePerCarat, aes(x=clarity, y=meanPrice))+
  geom_bar(stat="identity", fill = "red")+
  theme(plot.title = element_text(hjust = .5))+
  labs(title = "Mean Price per Carat by Clarity", 
       x = "Clarity",
       y = "Mean Price Per Carat")

ggplot(ColorVSPricePerCarat, aes(x=color, y=meanPrice))+
  geom_bar(stat="identity", fill = "blue")+
  theme(plot.title = element_text(hjust = .5))+
  labs(title = "Mean Price per Carat by Color", 
       x = "Color",
       y = "Mean Price Per Carat")

ggplot(CutVSPricePerCarat, aes(x=cut, y=meanPrice))+
  geom_bar(stat="identity", fill = "green")+
  theme(plot.title = element_text(hjust = .5))+
  labs(title = "Mean Price per Carat by Cut",
       x = "Cut",
       y = "Mean Price Per Carat")

```



#looking at some outliers
```{r}
diamonds%>%
  filter(clarity == "FL")

diamonds%>%
  filter(cut == "Ideal")

diamonds%>%
  filter(color == "H")
```



```{r}
ggplot(diamonds, aes(x=cut,fill=color))+
  geom_bar(position='fill')+
  theme(axis.text.x = element_text(),
        plot.title = element_text(hjust = 0.5))+
  labs(title="Distribution of cut vs color")
```
#Other than Astor Ideal (which only has 20 datapoints), good color seems to decrease as cut gets better. This corrlation could be a factor in why cut does not seem to correlate as much with price/carat.






# Regression Code








```{r}
library(MASS) 
library(tidyverse)
```

```{r}
dimonds <- read.csv("diamonds4.csv")
dimonds
```
```{r}
ggplot(data = dimonds, mapping = aes(x=carat, y = price))+
  geom_point()+
  labs(x="Carat", y="Price", title="General Scatter Plot of Price vs Carrat")
```


```{r}
ggplot(data = dimonds, mapping = aes(x=carat, y = price))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+ 
  labs(x="Carat", y="Price", title="General Scatter Plot of Price vs Carrat (With Regression Line)")
```


```{r}
Data = dimonds
result<-lm(price~carat, data=Data)
yhat<-result$fitted.values
res<-result$residuals
Data<-data.frame(Data,yhat,res)

## adding inital attributes to the Data DF
```



```{r}
ggplot(Data, aes(x=yhat,y=res))+
  geom_point()+
  geom_hline(yintercept=0, color="red")+
  labs(x="Fitted y", y="Residuals", title="Inital Residual Plot (No Transformations)")
```

```{r}
boxcox(result, lambda = seq(0,.5,1/10), main= "Box Cox (No Transformations")
```


```{r}
### First transformation:


##transform y and then regress ystar on x
ystar<-log(Data$price)
Data<-data.frame(Data,ystar)
result.ystar<-lm(ystar~carat, data=Data)

##store fitted y & residuals
yhat2<-result.ystar$fitted.values
res2<-result.ystar$residuals

##add to data frame
Data<-data.frame(Data,yhat2,res2)

##residual plot with ystar
ggplot(Data, aes(x=yhat2,y=res2))+
  geom_point()+
  geom_hline(yintercept=0, color="red")+
  labs(x="Fitted y", y="Residuals", title="Residual Plot with ystar")

boxcox(result.ystar, lambda = seq(2,4,1/10))
## acf plot
# par(mar=c(5,5,5,0))
# acf(res2, main="ACF Plot of Residuals with ystar")
# ## QQ plot
# qqnorm(res2)
# qqline(res2, col="red")


result.ystar
```

```{r}
### Second transformation:


##transform y and then regress ystar on x
xstar<-log(Data$carat)
Data<-data.frame(Data,xstar)
result.xstar<-lm(ystar~xstar, data=Data)

##store fitted y & residuals
yhat3<-result.xstar$fitted.values
res3<-result.xstar$residuals

##add to data frame
Data<-data.frame(Data,yhat3,res3)

##residual plot with ystar
ggplot(Data, aes(x=yhat3,y=res3))+
  geom_point()+
  geom_hline(yintercept=0, color="red")+
  labs(x="Fitted y", y="Residuals", title="Residual Plot with ystar")

boxcox(result.xstar)
## acf plot
par(mar=c(5,5,5,0))
acf(res3, main="ACF Plot of Residuals with ystar and xstar",lag.max = 15)
## QQ plot
qqnorm(res3)
qqline(res3, col="red")

result.xstar
```



```{r}
ggplot(data=Data, mapping = aes(x=xstar, y=ystar))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs(title="Final Regression with xstar and ystar")
  
```

```{r}
result.xstar
```

log(y) = 8.521 + 1.944 log(x)

#
